Luminance constancy under fixed geometry

### 2018-02-06 Draft2/draft2.tex
added more figures
more edits

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of object luminance under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived luminance of an object. We have developed a software that generates naturalistic multispectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the luminance of a specific object in the scene. Next, we simulate the response of the retinal cones to these multispectral images using an accurate model of the early visual system.  Using supervised learning methods on these labeled cone response, we identify the computations that lead to accurate luminance estimation. We show that if only either target, or background, or illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the luminance within about 15\% RMSE.}

@@

There is a considerable literature on computational color constancy. Seminal work considered a reduced scene model in which flat matte surfaces were illuminated by a single spatially diffuse illuminant and explored algorithms that could either recover a descriptor correlated with object surface reflectance (e.g. \citeNP{LandRetinex}) or explicit estimates of object surface reflectance and/or illuminant spectral power distribution (e.g., \citeNP{LandRetinex,Buchsbaum80,MaloneyWandell86}).
An important insight of this line of research was that, to account for variation in illumination, it was necessary to consider the light reflected from the entire scene. That is, although good constancy cannot be achieved in the face of illuminant variation if one considers only the light reflected from a single target object, when one considers the entire image it is possible to use the joint variation in light reflected from many objects to obtain (either implicitly or explicitly) an estimate of the illuminant common across the objects, and in turn to use this estimate to produce a stabilized representation of object color.
At the same time, this processing introduces an additional dependence.  Although the reflectance of the objects surrounding an object of interest has little direct effect on the light reflected from the object of interest, variation in the reflectance of the surrounding objects can have a large effect on the output of constancy algorithms that process the entire image \cite{BrainardWandellRetinex}.
Thus, a key aspect evaluating computational constancy algorithms is not only to ask about target object color stability with respect to variation in illumination, but also to ask about target object color stability with respect to the surface reflectance of other objects in the scene.
Subsequent computational constancy work sought to generalize both the image formation model to include richer geometric descriptions and to incorporate probabilistic descriptions of the statistics of naturally occurring scenes \cite{funt1988color, D'ZmuraConstancy3, barron2012color, D'ZmuraIversonSinger,BrainardFreeman}; LET'S SEE IF WE CAN FIND SOME RECENT REVIEWS).  None-the-less, it is not known how best to estimate stable object descriptors from images of the natural scenes we view in daily life.

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.45 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.jpeg}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light reflected from an object to our eyes depends both on the surface reflectance of the object, $R(\lambda,\vec{x})$, and the illumination spectrum of the light source, $E(\lambda,\vec{x})$. Additionally, it depends on geometric factors, such as the object's shape and pose, as well as the position of the observer. The human visual system is able to account for, at least partially, variations in the reflected light due to such extrinsic factors, and produce a percept of object color that is relatively stable with respect to such variations. Such invariant color perception is called color constancy. (b) {\bf Luminance Constancy Under Spectral Variations:} In this paper we consider the special case of luminance constancy. Fig.~\ref{fig:introExampleFigure} shows images of an object (the sphere in the center of each image) at three levels of object luminance (indicated on the left). Within each row, the surface reflectance of the spheres has the same luminance, when illuminated by a standard reference illuminant, CIE D65. The three rows differ in this value, as can be seen by the fact that the luminance of the spheres generally decreases as we move down the columns. Within each row, the relative surface reflectance of the spheres varies, as can be seen by the variation in color appearance across the each row.  Within each column, this relative surface reflectance is held fixed. For illustration, we have kept the light source and the background objects fixed across the panels. See Fig.~\ref{fig:studiedCases} for images with such variations. In this paper, we seek to identify computations that accurately estimate object luminance across variations in relative surface reflectance, variations in illumination, and variation in the reflectance of other objects in the scene.}
 \end{figure}

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate ``inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron \cite{barron2015convolutional} recently developed a color balancing algorithm based on training a neural network to detect a uniform illuminant in an image by matching image templates in the chromaticity space. The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision. Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, \citeNP{geisler2009optimal}), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate speed \cite{burge2015optimal}, focus \cite{burge2011optimal} and disparity \cite{burge2014optimal} from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA, as well as comparison supervised learning algorithms to a special case of the color constancy problem (luminance constancy) and evaluate how well it performs for naturalistic image input.

@@

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 2}
 		\label{fig:backgroundVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 3}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 4}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Three types of spectral variations studied in this work:} sRGB rendition of multispectral images similar to ones studied in our work. The numbers on the left indicate the standard luminance level of the target object. 5 images are shown at each luminance level. We studied four types of spectral variations. Case 1: target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra fixed, (Fig.~\ref{fig:introExampleFigure}). (a) Case 2: target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra variable, (b) Case 3: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra fixed, (c) Case 4: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra variable. As in \ref{fig:introExampleFigure}, the spheres in each row of each panel have the same surface luminance, while the spheres in each column of each panel have the same relative surface reflectance.  Moreover, across the three panels of this figure, spheres in corresponding locations have the same surface reflectance. In all three panels, the overall lights source spectra scale factors (see \nameref{Methods}) were drawn from a uniform distribution on the range [0.5, 1.5]. The sRGB values for all three panels were normalized using a common scale factor prior to gamma correction.} 
\label{fig:studiedCases}
\end{figure}

@@

The rest of this paper is organized as follows. In the \nameref{Methods} section, we describe in detail the process to generate naturalistic labeled images (\nameref{method:VirtualWorld}) using our software pipeline {\rm Virtual world color constancy}, the process to simulate the response of retinal cone photoreceptors to these images using a standard model of the early visual system (\nameref{method:Isetbio}) and the supervised learning method (AMA) used in our analysis (\nameref{method:SupervisedLearning}). In the \nameref{Results} section, we show the estimates and performance of the supervised methods for the three cases mentioned above. We conclude the paper with the implication of our results in the \nameref{Discussion} section.

\section*{Methods} \label{Methods}
\subsection{Overview}
There are four key parts to our methods.  The first is how we generate a labelled set of training data for our supervised learning algorithms.  The second is how we model the transformations applied by the early visual system to generate modeled responses of the cone photoreceptor mosaic.  The third is the details of the supervised learning method (AMA) we apply to the labelled training data.  And the fourth is how we evaluate the performance of the algorithms.  We describe each of these in the sections below.

@@

% Figure 3
\begin{figure}[t]
\centering
\begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_a.png}
        \caption{Library }
        \label{fig:baseSceneLibrary}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_b.png}
        \caption{Mill}
        \label{fig:baseSceneMill}
    \end{subfigure}    
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_c.png}
        \caption{Table-Chairs}
        \label{fig:baseSceneTableChairs}
    \end{subfigure}

@@

% Figure 4
\begin{figure}
\centering
\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_a.png}
        \caption{Big-ball}
        \label{fig:libraryWithBigBall}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_b.png}
        \caption{Small-ball}
        \label{fig:libraryWithSmallBall}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_c.png}
        \caption{Barrel}
        \label{fig:libraryWithBarrel}
    \end{subfigure}
    \begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_d.png}
        \caption{Xylophone}
        \label{fig:libraryWithXylophone}
    \end{subfigure}
    ~
	\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_e.png}
        \caption{Ring toy}
        \label{fig:libraryWithRingToy}
    \end{subfigure}
        ~
    	\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_f.png}
        \caption{Bottle}
        \label{fig:libraryWithChampagneBottle}
    \end{subfigure}
\caption{{\bf Library base scene with inserted objects.} The VWCC software was used to insert different objects into the library base scene, with each panel showing a different object. The objects were inserted at a location in the image and then the camera was pointed at the object, so that the object's image is at the center of the rendered image.  In the figure panels, the full rendered image is cropped so that the inserted object is more visually salient. We use this capability of our software pipeline to insert target objects into scenes.}\label{fig:libraryWithTarget}
\end{figure}

The process of generating an individual virtual scene begins with the selection of a \textit{base scene} (Fig.~\ref{fig:baseScenes}). The base scene is a 3D model that defines a rendering space.  It typically includes the specification of a number of 3D objects and light sources, and is annotated with meta-data that may be used to guide the placement of additional objects and light sources, as well as the specification of camera position. VWCC includes a library of base scenes obtained from materials made available on the Internet under various Creative Commons licenses. We have found that there is sufficient variation in how nominally-standard graphics interchange file formats are used in practice that each base scene obtained from the Internet (which we refer to as \textit{wild scenes}) requires a certain amount of hand editing before it renders properly. We refer to this process, along with adding meta-data about the scene, as \textit{taming} a wild scene. Tame base scenes may also be created de novo using a 3D modeling software (e.g., \href{https://www.blender.org/}{Blender}).  The overall spatial scale of the base scene may be adjusted under programmatic control or chosen randomly, adding to the available statistical richness. If one chooses to draw entries from the base scene library at random for each virtual scene, the base scene library then provides a simple statistical model of the gross layout of natural scenes. As time progresses, we plan to tame more wild scenes so as to increase the number of entries in the base scene library. Currently the library contains 6 base scenes: Library (Fig.~\ref{fig:baseSceneLibrary}), Mill (Fig.~\ref{fig:baseSceneMill}), Table-Chairs (Fig.~\ref{fig:baseSceneTableChairs}), Indoor plant (Fig.~\ref{fig:baseSceneIndoorPlant}), Checkerboard (Fig.~\ref{fig:baseSceneWarehouse}) and Warehouse (Fig.~\ref{fig:baseSceneCheckerBoard}).

@@

% Figure5
\begin{figure}
\centering
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_a.png}
        \caption{Target spectra}
        \label{fig:targetVariation}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_b.png}
        \caption{Background spectra}
        \label{fig:backGroundVariation}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_c.png}
        \caption{Illumination spectra}
        \label{fig:illuminationVariation}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_d.png}
        \caption{Target position}
        \label{fig:targetPositionVariation}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_e.png}
        \caption{Target Size/Orientation}
        \label{fig:targetSizeOrientation}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \caption{{\bf Transformations of the scene.} The possible transformations of the properties of a scene can broadly be classified into two groups: Spectral (a-c) and Geometrical (d-e). VWCC provides control over such transformation as illustrated by the columns. (a) The target object surface reflectance varies across the three panels of the column. (b) The surface reflectance of the background objects varies across the three panels. In each panel, the reflectances were assigned randomly. (c) The power spectrum of the light sources were assigned randomly in the three panels. (d) The three panels in the column have the target object at three different positions. The camera points to the center of the pixel. }\label{fig:VWCCTransformations}
\end{figure}

@@

For the images used for the primary results reported in this paper, we used the ``{\it Library}'' base scene from the VWCC  and the ``{\it Big Ball}'' object as the target object. We also inserted a spherical light source with the same ``{\it Big Ball}'' shape. The position, and size of the inserted object, light source and the camera were held fixed for all the images analyzed in this work, i.e., the scene geometry did not vary. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work have been described in the main text (Fig.~\ref{fig:studiedCases}). 2D multispectral images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel. A $51 \times 51$ pixel patch of the multispectral image was cropped around the target object and cone responses were calculated to these cropped images.

@@

The isomerizations of the cones in the retinal mosaic represent the first steps of visual processing.  To capture key properties of post-receptoral processing, contrast coding (REFS) and contrast normalization (REFS),
% Ask Johannes what he'd use as references for these broad principles
we processed the isomerizations as follows:
First, we created a representation for each cone class on the full $51 \times 51$ pixel grid, we interpolated the responses of the cones in each class across the grid, using bilinear interpolation separately for each cone class.
% DOUBLE CHECK THAT THIS WAS DONE WITH SOMETHING LIKE INTERP2 WITH THE LINEAR OPTION.
This gives us a $51 \times 51$ cone response image for each cone class.
The three $51 \times 51$ sets of cone responses corresponding to the L, M and S cones were then reshaped into a single $7803 \times 1$ vector ($7803 = 51 \times 51 \times 3$). To model the contrast coding and contrast adaptation, we convert this vector to a contrast representation by subtracting out the mean and then dividing by it (contrast coding), followed by normalizing the vector to have unit length (contrast normalization).
The cone-response vectors corresponding to the N images for each case are concatenated column-wise to form a contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.

>>>>

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of object luminance under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived luminance of an object. We have developed a software that generates naturalistic multispectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the luminance of a specific object in the scene. Next, we simulate the response of the retinal cones to these multispectral images using an accurate model of the early visual system. We then use supervised learning methods on these labeled cone response to identify the computations that lead to accurate luminance estimation. We show that if only either target, or background, or illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the luminance within about 15\% RMSE.}

@@

There is a considerable literature on computational color constancy. Seminal work considered a reduced scene model in which flat matte surfaces were illuminated by a single spatially diffuse illuminant and explored algorithms that could either recover a descriptor correlated with object surface reflectance (e.g. \citeNP{LandRetinex}) or produce explicit estimates of object surface reflectance and/or illuminant spectral power distribution (e.g., \citeNP{LandRetinex,Buchsbaum80,MaloneyWandell86}).
An important insight of this line of research was that, to account for variation in illumination, it was necessary to consider the light reflected from the entire scene. That is, although good constancy cannot be achieved in the face of illuminant variation if one considers only the light reflected from a single target object, when one considers the entire image it is possible to use the joint variation in light reflected from many objects to obtain (either implicitly or explicitly) an estimate of the illuminant common across the objects, and in turn to use this estimate to produce a stabilized representation of object color.
At the same time, this processing introduces an additional dependence.  Although the reflectance of the objects surrounding an object of interest has little direct effect on the light reflected from the object of interest, variation in the reflectance of the surrounding objects can have a large effect on the output of constancy algorithms that process the entire image \cite{BrainardWandellRetinex}.
Thus, a key aspect evaluating computational constancy algorithms is not only to ask about target object color stability with respect to variation in illumination, but also to ask about target object color stability with respect to the surface reflectance of other objects in the scene.
Subsequent computational constancy work sought to generalize both the image formation model to include richer geometric descriptions and to incorporate probabilistic descriptions of the statistics of naturally occurring scenes \cite{funt1988color, D'ZmuraConstancy3, barron2012color, D'ZmuraIversonSinger,BrainardFreeman}; LET'S SEE IF WE CAN FIND SOME RECENT REVIEWS).  None-the-less, it is not known how best to estimate stable object descriptors from images of the natural scenes we view in daily life.

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.45 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.jpeg}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light reflected from an object to our eyes depends both on the surface reflectance of the object, $R(\lambda,\vec{x})$, and the illumination spectrum of the light source, $E(\lambda,\vec{x})$. Additionally, it depends on geometric factors, such as the object's shape and pose, as well as the position of the observer. The human visual system is able to account for, at least partially, variations in the reflected light due to such extrinsic factors, and produce a percept of object color that is relatively stable with respect to such variations. Such invariant color perception is called color constancy. (b) {\bf Luminance Constancy Under Spectral Variations:} In this paper we consider the special case of luminance constancy. Fig.~\ref{fig:introExampleFigure} shows images of an object (the sphere in the center of each image) at three levels of object luminance (indicated on the left). Within each row, the surface reflectance of the spheres has the same luminance when illuminated by a standard reference illuminant, CIE D65. The three rows differ in this value, as can be seen by the fact that the luminance of the spheres generally decreases as we move down the columns. Within each row, the relative surface reflectance of the spheres varies, as can be seen by the variation in color appearance across the each row.  Within each column, this relative surface reflectance is held fixed. For illustration, we have kept the light source and the background objects fixed across the panels. See Fig.~\ref{fig:studiedCases} for images with such variations. In this paper, we seek to identify computations that accurately estimate object luminance across variations in relative surface reflectance, variations in illumination, and variation in the reflectance of other objects in the scene.}
 \end{figure}

The computational work reviewed above is based on formulating an explicit model of the image formation process and the statistical properties of natural images, and leveraging these formulations to obtain an appropriate ``inverse optics" algorithm.  An alternative approach, which has been less-well explored, is to use supervised machine learning techniques to develop mappings between input images and stable color descriptors.  Along these lines, Barron \cite{barron2015convolutional} recently developed a color balancing algorithm based on training a neural network to detect a uniform illuminant in an image by matching image templates in the chromaticity space. The use of supervised learning to understand human perceptual capabilities has enjoyed success in domains outside of color vision. Geisler and colleagues, for example, have developed a technique they refer to as accuracy maximization analysis (AMA, \citeNP{geisler2009optimal}), which learns linear filters whose output carries information that is optimized for particular perceptual tasks. This technique has provided computations that efficiently estimate speed \cite{burge2015optimal}, focus \cite{burge2011optimal} and disparity \cite{burge2014optimal} 
% JOHANNES ARE THERE ANY OTHER PAPERS YOU WOULD LIKE TO CITE
from image patches taken from natural scenes, and that in addition provide excellent models of human's ability to make corresponding estimates from the same stimuli.  In this paper, we apply AMA, as well as comparison supervised learning algorithms to a special case of the color constancy problem (luminance constancy) and evaluate how well it performs for naturalistic image input.

@@

% Figure 2: Cases Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_a.jpeg}
		\caption{Case 2}
 		\label{fig:backgroundVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_b.jpeg}
        \caption{Case 3}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure2/Figure2_c.jpeg}
        \caption{Case 4}
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Types of spectral variations studied in this work:} sRGB rendition of multispectral images similar to ones studied in our work. The numbers on the left indicate the standard luminance level of the target object. 5 images are shown at each luminance level. We studied four types of spectral variations. Case 1: target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra fixed, (Fig.~\ref{fig:introExampleFigure}). (a) Case 2: target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra variable, (b) Case 3: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra fixed, (c) Case 4: target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra variable. As in \ref{fig:introExampleFigure}, the spheres in each row of each panel have the same surface luminance, while the spheres in each column of each panel have the same relative surface reflectance.  Moreover, across the four panels (Fig.~\ref{fig:introExampleFigure} \& Fig.~\ref{fig:studiedCases}), spheres in corresponding locations have the same surface reflectance. In all four panels, the overall lights source spectra scale factors (see \nameref{Methods}) were drawn from a uniform distribution on the range [0.5, 1.5]. The sRGB values for all three panels were normalized using a common scale factor prior to gamma correction.} 
\label{fig:studiedCases}
\end{figure}

@@

The rest of this paper is organized as follows. In the \nameref{Methods} section, we describe in detail the process to generate naturalistic labeled images (\nameref{method:VirtualWorld}) using our software pipeline {\rm Virtual world color constancy}, the process to simulate the response of retinal cone photoreceptors to these images using a standard model of the early visual system (\nameref{method:Isetbio}) and the supervised learning method (AMA) used in our analysis (\nameref{method:SupervisedLearning}). In the \nameref{Results} section, we show the estimates and performance of the supervised methods for the four cases mentioned above. We conclude the paper with the implication of our results in the \nameref{Discussion} section.

\section*{Methods} \label{Methods}
\subsection{Overview}
There are four key parts to our methods.  The first is how we generate a labelled set of training data for our supervised learning algorithms.  The second is how we model the transformations applied by the early visual system to generate modeled responses of the cone photoreceptor mosaic.  The third is the details of the supervised learning method (AMA) that we apply to the labelled training data.  And the fourth is how we evaluate the performance of the algorithm.  We describe each of these in the sections below.

@@

% Figure 3
\begin{figure}[t]
\centering
\begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_a.png}
        \caption{Library }
        \label{fig:baseSceneLibrary}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_b.png}
        \caption{Mill}
        \label{fig:baseSceneMill}
    \end{subfigure}    
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure3/Figure3_c.png}
        \caption{Table-Chairs}
        \label{fig:baseSceneTableChairs}
    \end{subfigure}

@@

% Figure 4
\begin{figure}
\centering
\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_a.png}
        \caption{Big-ball}
        \label{fig:libraryWithBigBall}
    \end{subfigure}
    ~ 
\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_b.png}
        \caption{Small-ball}
        \label{fig:libraryWithSmallBall}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_c.png}
        \caption{Barrel}
        \label{fig:libraryWithBarrel}
    \end{subfigure}
    \begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_d.png}
        \caption{Xylophone}
        \label{fig:libraryWithXylophone}
    \end{subfigure}
    ~
	\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_e.png}
        \caption{Ring toy}
        \label{fig:libraryWithRingToy}
    \end{subfigure}
        ~
    	\begin{subfigure}[b]{0.14 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure4/Figure4_f.png}
        \caption{Bottle}
        \label{fig:libraryWithChampagneBottle}
    \end{subfigure}
\caption{{\bf Library base scene with inserted objects.} The VWCC software was used to insert different objects into the library base scene, with each panel showing a different object. The objects were inserted at a location in the image and then the camera was pointed at the object, so that the object's image is at the center of the rendered image.  In the figure panels, the full rendered image is cropped so that the inserted object is more visually salient. We use this capability of our software pipeline to insert target objects into scenes.}\label{fig:libraryWithTarget}
\end{figure}

The process of generating an individual virtual scene begins with the selection of a \textit{base scene} (Fig.~\ref{fig:baseScenes}). The base scene is a 3D model that defines a rendering space.  It typically includes the specification of a number of 3D objects and light sources, and is annotated with meta-data that may be used to guide the placement of additional objects and light sources, as well as the specification of camera position. VWCC includes a library of base scenes obtained from materials made available on the Internet under various Creative Commons licenses. We have found that there is sufficient variation in how nominally-standard graphics interchange file formats are used in practice that each base scene obtained from the Internet (which we refer to as \textit{wild scenes}) requires a certain amount of hand editing before it renders properly. We refer to this process, along with adding meta-data about the scene, as \textit{taming} a wild scene. Tame base scenes may also be created {\it de novo} using a 3D modeling software (e.g., \href{https://www.blender.org/}{Blender}).  The overall spatial scale of the base scene may be adjusted under programmatic control or chosen randomly, adding to the available statistical richness. If one chooses to draw entries from the base scene library at random for each virtual scene, the base scene library then provides a simple statistical model of the gross layout of natural scenes. As time progresses, we plan to tame more wild scenes so as to increase the number of entries in the base scene library. Currently the library contains 6 base scenes: Library (Fig.~\ref{fig:baseSceneLibrary}), Mill (Fig.~\ref{fig:baseSceneMill}), Table-Chairs (Fig.~\ref{fig:baseSceneTableChairs}), Indoor plant (Fig.~\ref{fig:baseSceneIndoorPlant}), Checkerboard (Fig.~\ref{fig:baseSceneWarehouse}) and Warehouse (Fig.~\ref{fig:baseSceneCheckerBoard}).

@@

% Figure5
\begin{figure}
\centering
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_a.png}
        \caption{Target spectra}
        \label{fig:targetVariation}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_b.png}
        \caption{Background spectra}
        \label{fig:backGroundVariation}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_c.png}
        \caption{Illumination spectra}
        \label{fig:illuminationVariation}
    \end{subfigure}
    ~
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_d.png}
        \caption{Target position}
        \label{fig:targetPositionVariation}
    \end{subfigure}
    ~
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure5/Figure5_e.png}
        \caption{Target Size/Orientation}
        \label{fig:targetSizeOrientation}
    \end{subfigure}
    ~
    \caption{{\bf Transformations of the scene.} The possible transformations of the properties of a scene can broadly be classified into two groups: Spectral (a-c) and Geometrical (d-e). VWCC provides control over such transformation as illustrated by the columns. (a) The target object surface reflectance varies across the three panels of the column. (b) The surface reflectance of the background objects varies across the three panels. In each panel, the reflectances were assigned randomly. (c) The power spectrum of the light sources were assigned randomly in the three panels. (d) The three panels in the column have the target object at three different positions. The camera points to the center of the pixel. }\label{fig:VWCCTransformations}
\end{figure}

@@

For the images used for the primary results reported in this paper, we used the ``{\it Library}'' base scene from the VWCC  and the ``{\it Big Ball}'' object as the target object. We also inserted a spherical light source with the same ``{\it Big Ball}'' shape. The position, and size of the inserted object, light source and the camera were held fixed for all the images analyzed in this work, i.e., the scene geometry did not vary. Next, we assigned illumination and surface reflectance spectra to the lights and objects in the scene. The spectral manipulations for the various cases studied in this work have been described in the main text (Fig.~\ref{fig:introExampleFigure} \& Fig.~\ref{fig:studiedCases}). 2D multispectral images were rendered at 31 wavelengths from $400$nm to $700$nm spaced $10$nm apart at an image size $320\times 240$ pixel. A $51 \times 51$ pixel patch of the multispectral image was cropped around the target object and cone responses were calculated to these cropped images.

@@

The isomerizations of the cones in the retinal mosaic represent the first steps of visual processing.  To capture key properties of post-receptoral processing, contrast coding (REFS) and contrast normalization (REFS),
% Ask Johannes what he'd use as references for these broad principles
we processed the isomerizations as follows:
First, we created a representation for each cone class on the full $51 \times 51$ pixel grid, we interpolated the responses of the cones in each class across the grid, using bilinear interpolation separately for each cone class.
% DOUBLE CHECK THAT THIS WAS DONE WITH SOMETHING LIKE INTERP2 WITH THE LINEAR OPTION.
This gives us a $51 \times 51$ cone response image for each cone class.
The three $51 \times 51$ sets of cone responses corresponding to the L, M and S cones were then reshaped into a single $7803 \times 1$ vector ($7803 = 51 \times 51 \times 3$). For case 1 and 3 we used this as the input for the supervised methods. For case 3 and 4, to model the contrast coding and contrast adaptation, we convert this vector to a contrast representation by subtracting out the mean and then dividing by it (contrast coding), followed by normalizing the vector to have unit length (contrast normalization).
The cone-response vectors corresponding to the N images for each case are concatenated column-wise to form a contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.
